%% Why don't we use move; Update-detector algorithm
\section{A solution for update-detection for a move--free algebra}
\label{app:update-move}

\begin{comment}
\subsection{A simple way}
\remark{Do we need this in the paper?}
A ``simple way'' means that when determining the commands which lead from
the
original to the present state of the filesystems we do not use commands
like moving a whole subtree or removing a subtree. Commands only allowed
at leaves of the tree of the filesystem, e.g. files or directories
containing no files.

Consider the original state, called \(O\) and the present state called
\(A\). In order to be able to determine the \emph{move} operations, we'd
like to compare the containings of two files in the original and in the
present state. It doesn't mean that we have to store the original
filesystem. This can be achieved by a method called
\emph{data--fingerprint} with which we can compare contents of files
without comparing the whole amount of data stored in them. 

A function called \(origin(\pi)\) gives
us a path for a file in the filesystem \(O\) which has the same contents 
as the file \(A\cdot\pi\), or empty--file (\(\perp\)) if such a file does
not exist. Note that more files may be found in \(O\) with the same
contents.

Another function, \(\#(\pi)\) gives the number of paths \(\varphi\) for
which \(origin(\varphi)=\pi\) holds.

Then to gain a possible sequence of commands, compare \(O\cdot\pi\) and
\(A\cdot\pi\). (For this purpose, we can use timestamps, inode numbers,
or any other data on files provided by the operation system.) 
If they're the same, we
don't need to do anything. If not, we have three cases. First, if the
file \(A\cdot\pi\) has been deleted, we can delete this file iff
\(\#(\pi)=0\). If the file exists in \(A\) it may have an origin or not.
If it has, and \(\#(\pi)=0\), than we can \emph{move} \(origin(\pi)\) to
\(\pi\) iff \(\#(origin(\pi))=1\), or \emph{copy} it if
\(\#(origin(\pi))>1\). In both cases, we should decrease
\(\#(origin(\pi))\) by one.

If we couldn't do anything because \(\#(\pi)\) was greater than \(0\), we
can do one of the followings: wait till it decreases due to another
commands, or, if there's a \emph{loop} of file displacings (for example
\(origin(a)=b\) and \(origin(b)=a\)) than we should rename the origin of
one of the files in order to be able to continue (solving the
example: \emph{move} \(b\to b^\prime\), \(a\to b\) and \(b^\prime\to
a\)).

That way we gained a possible sequence of commands which if we apply
to 
filesystem \(O\) we'll get filesystem \(A\).
\end{comment}

First, we provide an example which explains why we chose working on
move--free algebra.

\subsection{Moving and removing subtrees}

When trying to determine the minimal sequence of commands which was
applied to the original filesystem \(O\) so that we get its present state
\(A\) we should try to move
or remove subtrees instead of (re)moving all files in it
in order to gain the most simple sequence possible.
We can achieve
this by counting the files under a directory and how many of
them should be (re)moved. If there are many files to be moved, we should
move the subtree first and then the other files back to their place
(majority voting).

\subsection{Problems with move}

But, unfortunately, difficulties emerge when using such commands.

First of all, consider the following situation: we have files \(X\cdot1,
X\cdot2, X\cdot3, X\cdot4, X\cdot5, X\cdot6\) in filesystem \(O\). In
replica A, we have moved \(X\cdot1, \ldots X\cdot4\) to \(Y\cdot1, \ldots
Y\cdot4\). In replica B, we have moved only \(X\cdot1, X\cdot2\) under
the new directory \(Y\).

In replica \(A\), without using commands on subtrees, we'd
have \emph{move} \(X\cdot1\to Y\cdot1, X\cdot2\to Y\cdot2,
X\cdot3\to Y\cdot3, X\cdot4\to Y\cdot4\). But using subtree--moving,
we'll
notice that using \emph{move} \(X\to Y, Y\cdot5\to X\cdot5,
Y\cdot6\to X\cdot6\) would be shorter, so this is our sequence of
commands. In replica \(B\), \emph{move}
\(X\cdot1\to Y\cdot1, X\cdot2\to Y\cdot2\) is the shortest way.

Now we want to choose commands from each sequence so that they would not
cause a conflicting update if propagated on other replicas.
It is easily provable that such subsequences of commands
do not exist. 
\remark{Proof?}
We reached the conclusion that not always the
shortest sequence should be used to generate the interleaving; or, in
other words, we should be able to change the sequences according to the
algebraic laws before choosing. It'd make the problem much more
complicated since we have a lot of possibilities here as well as at
choosing the subsequences.  Using \(move\), we also have a very 
complicated relationship among the commands considering which two can
commute and which two not. (Note that \(move\) is the only function
which modifies \emph{two} paths at the same time.)

%We can solve this problem by \emph{delaying
%the subtree--detection} to the end of the update--detection and
%reconciliation.

Therefore, we decided to focus on a \emph{move--free} algebra and delay
detecting possibilities of simplification (subtree--detection) until
after the update--detection and reconciliation.

\subsection{A simple algorithm for update--detection}
\label{app:upd}

With no \(move\)s and subtree--commands, a simple algorithm can be
used for update--detection. It takes information from the filesystems
and gives a minimal sequence of commands which will make one similar to
the other. 

Let \(P_O\) be all the paths in the original
filesystem, \(P_A\) all the paths in the current state of the replica. 
(We assume \(O,A\ne\ud\).)
For
every \(\pi\in P_O\cup P_A\) add the following command to sequence \(S\):
\begin{itemize}
\item \(create(\pi,X)\) iff \(O(\pi)=\ud\) and \(A(\pi)=X\ne\ud\),
\item \(edit(\pi,X)\) iff \(O(\pi)\ne\ud\) and \(A(\pi)=X\ne\ud\) and
\(X\ne O(\pi)\),
\item \(remove(\pi)\) iff \(O(\pi)\ne\ud\) and \(A(\pi)=\ud\).
\end{itemize}
Then arrange the commands so that
\begin{itemize}
\item all \(edit(\pi,\cdir{X})\) commands precede all other commands, 
\item all \(edit(\pi,\cfile{X})\) commands follow all other commands,
\item \(create(\pi,X)\) commands precede \(create(\pisub,Y)\) commands,
\item \(remove(\pisub,X)\) commands precede \(remove(\pi,Y)\) commands.
\end{itemize}
This can be done by the method discussed below.

First, we move all \(edit(\pi,\cdir{X})\) commands to the front and then
all \(edit(\pi,\cfile{X})\) commands to the end of the sequence. Now for
the middle part, we know the following:

Since there is only one command on each path, we have a partial
ordering on the set of the commands (see section
\ref{theorem:conc}) showing us which command must precede another one.
Two commands can be interchanged if they refer to incomparable
paths; otherwise, they are one of the following:
\begin{itemize}
\item \(create(\pi) \ldots create(\pisub)\)
\item \(remove(\pisub) \ldots remove(\pi)\).
\end{itemize}
That is, every \(create\) must precede all other \(create\)s on its
descendants and all \(remove\)s must precede \(remove\)s on its parents.

This ordering can be represented by a graph. This graph is a tree (or
forest). It's worth remarking that every tree of the forest is formed
from the same commands, \(remove\)s or \(create\)s, since two different
types of commands cannot show up at comparable paths. 

Starting at the leaves of the trees, we move the leaves as close as
possible to their parent. (They will form a row at the left side of the
branch--command at the \(remove\)--trees and at the other side at
\(create\)--trees.) Then we order them alphabetically and form a
\emph{group--command} from the leaves and the branch. This
group--command
will act like a single command since all of its members commute the
same commands. Then we continue this method until every tree is
grouped into one meta-group--command. We order these groups again.
That way we gain an order of the commands. It's worth remarking that
we gain the same order for the same set of commands every time.

We know that \(SO=A\) since because of the ordering it does not break the
filesystem. Also, \(S\) is a minimal sequence. \emph{Proof.} 
By contradiction.
If \(S'\) is shorter than \(S\), then we have a path \(\varphi\) on which
we have a command in \(S\) but not in \(S'\) since all paths are different
in \(S\). Now we know that \(SF(\varphi)\ne F(\varphi)\) but 
\(S'F(\varphi)=F(\varphi)\). (Note that we have no superfluous commands
in \(S\) since we gained it from the update detector algorithm.)

