
% B 4
\bibitem{BP}
S. Balasubramaniam and Benjamin C. Pierce,
``What is a File Synchronizer?''
in \emph{Proceedings of the 4\textsuperscript{th} Annual 
ACM/IEEE International Conference on Mobile Computing and Networking.}
New York, NY: ACM, 1998, pp. 98--108.

% B 21
\bibitem{CBNR} 
Constantin Berzan and Norman Ramsey,
``Summer Scholars Technical Report,''
Tufts University, Medford, MA, 2010.
Accessed 29 November 2015.
http://thirld.com/files/summerscholars\undersc techreport.pdf.

% C 16
\bibitem{CH}
Yek Loong Chong and Youssef Hamadi,
``Distributed Algorithms for Log-based Reconciliation,''
Microsoft Research, Microsoft Co., Redmond, WA,
Tech. Rep. MSR-TR-2004-104, 2004.

% C 8
\bibitem{CJ}
Russ Cox and William Josephson,
``File synchronization with Vector Time Pairs,''
MIT CSAIL, Cambridge, MA, 
Tech. Rep., MIT-CSAIL-TR-2005-014, 2005.
Accessed 22 November 2015.
http://hdl.handle.net/1721.1/30527.

% F 18
\bibitem{FGKPS}
% Foster, J Nathan and Greenwald, Michael B and Kirkegaard, Christian and Pierce, Benjamin C and Schmitt, Alan
J. Nathan Foster et al.,
``Schema-directed data synchronization,''
U of Pennsylvania,
Philadelphia, PA,
Technical Reports (CIS),
Paper 5,
2005.
Accessed 23 November 2015.\\
http://repository.upenn.edu/cis\undersc reports/5.

% K 3
\bibitem{KRSD}
Anne-Marie Kermarrec and Antony Rowstron and Marc Shapiro and Peter Druschel,
``The IceCube Approach to the Reconciliation of Divergent Replicas,''
in \emph{Proceedings of the Twentieth Annual ACM Symposium on Principles of Distributed Computing.}
New York, NY: ACM, 2001, pp. 210--218.

% M 19
\bibitem{MPV}
Vidal Martins and Esther Pacitti and Patrick Valduriez,
``Distributed semantic reconciliation of replicated data,''
in \emph{Proc. of CDUR.}
Paris, France: CNAM,
2005,
pp. 48--53.

% P 6
\bibitem{PPRS}
% Parker, D. S. and Popek, G. J. and Rudisin, G. and Stoughton, A. and Walker, B. J. and Walton, E. and Chow, J. M. and Edwards, D. and Kiser, S. and Kline, C.
D. S. Parker et al.,
``Detection of Mutual Inconsistency in Distributed Systems,''
in \emph{IEEE Trans. Softw. Eng.,}
vol. 9, no. 3.
Piscataway, NJ: IEEE Press,
May 1983,
pp. 240--247.

% P 14
\bibitem{PV}
Benjamin C Pierce and J{\'e}r{\^o}me Vouillon,
``What's in Unison? A formal specification and reference implementation of a file synchronizer,''
U of Pennsylvania,
Philadelphia, PA,
Technical Reports (CIS),
Paper 40,
2004.
Accessed 23 November 2015.\\
http://repository.upenn.edu/cis\undersc reports/40.

% P 9
\bibitem{PSG}
Benjamin C. Pierce, Alan Schmitt, and Michael B. Greenwald, 
``Bringing Harmony to Optimism: An Experiment in Synchronizing Heterogeneous Tree-Structured Data,''
U of Pennsylvania,
Philadelphia, PA,
Technical Reports (CIS),
Paper 39,
2004.
Accessed 23 November 2015.\\
http://repository.upenn.edu/cis\undersc reports/39.

% R
\bibitem{NREC} 
Norman Ramsey and Elod Csirmaz,
``An Algebraic Approach to File Synchronization,''
in \emph{Proceedings of the Joint 8\textsuperscript{th} European Software Engineering Conference 
and 9\textsuperscript{th} ACM SIGSOFT Symposium on the Foundations of Software Engineering.}
New York, NY: ACM, 2001, pp. 175--185.

% S 12
\bibitem{SSH}
Yasushi Saito and Marc Shapiro,
``Optimistic Replication,''
in \emph{ACM Computing Surveys,}
vol. 37, no. 1.
New York, NY: ACM, March 2005,
pp. 42--81.

% T 1
\bibitem{TSR}
Vinh Tao and Marc Shapiro and Vianney Rancurel,
``Merging Semantics for Conflict Updates in Geo-distributed File Systems,''
in \emph{Proceedings of the 8\textsuperscript{th}
 ACM International Systems and Storage Conference.}
New York, NY: ACM, 2015, pp. 10:1--10:12.

% T 5
\bibitem{TTPDSH}
% D. B. Terry and M. M. Theimer and Karin Petersen and A. J. Demers and M. J. Spreitzer and C. H. Hauser
D. B. Terry et al.,
``Managing Update Conflicts in Bayou, a Weakly Connected Replicated Storage System,''
in \emph{Proceedings of the Fifteenth ACM Symposium on Operating Systems Principles.}
New York, NY: ACM, 1995, pp. 172--182.

% Z
\bibitem{BZ}
Bill Zissimopoulos, private communication, August 2015.
